---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Getting Started
### Prerequisites

Install packages library <b> palmerpenguins </b> and <b>tidymodels</b> 

```{r, eval = FALSE}
install.packages('palmerpenguins')
install.packages('tidymodels')
```

### Load package libraries
Load libraries
```{r}
library(dplyr)
library(palmerpenguins)
library(data.table)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(ggfortify)
library(recipes)
library(rsample)
library(yardstick)
library(caret)
library(profvis)
library(purrr)
```
## Load data
```{r}
penguins <- penguins %>% 
  mutate_if(is.character, as.factor) %>% 
  drop_na()
```

## EDA
The <b>penguins</b> dataset includes measurements of flow characteristics of the components. Only pterosaur data are considered in this project
```{r }
glimpse(penguins)
```
```{r }
summary(penguins)
```
```{r }
penguins %>% select_if(Negate(is.factor)) %>% 
  cor() %>% 
  as.data.frame() %>% 
  rowid_to_column() %>% 
  pivot_longer(- rowid) %>% 
  arrange(desc(abs(value))) %>% 
  filter(value != 1)
```
```{r, eval = FALSE }
site_number <- 1:4
pdf('penguin_data.pdf')
par(mfrow = c(2, 2))
for (i in site_number)
{
  penguins %>% 
    filter(site_number == i) %>% 
    ggplot(aes(x= species, y= flipper_length_mm, fill = species, color= species)) +
    geom_boxplot() +
    labs(x = 'Species', y = 'Flipper length (mm)', title = paste('Fvuwerfvczefcfzccfgct', i)) +
    theme(axis.text = element_text(size = 8))
}
dev.off()
```

```{r, eval = FALSE }
ggviolin(penguins, x = species,
        y = bill_length_mm+bill_depth_mm,
        fill = species)
```

```{r, eval = FALSE }
glimpse(penguins$sex)
table(penguins$sex, useNA='ifany')
```
## Data preparation
Set for training and testing data

```{r, eval = FALSE }
set.seed(3223)
train_test_split <- penguins %>% initial_split(prop = .7)
penguin_train <- training(train_test_split)
penguin_test <- testing(train_test_split)
glimpse(penguin_train)
```

Create recipe

```{r, eval = FALSE }
softmaxkitchen <- recipe(sex ~ ., data = penguin_train) %>%
 step_other(antipodean_island, all_nominal())%>%
step_novel(all_nominal())%>%
step_other(bill_length_mm, one_of('median-virginica'))%>%
step_other(bill_depth_mm, one_of('median-virginica'))
```

## Visualize recipe steps

```{r, eval = FALSE }
print(softmaxkitchen)
```

```{r, eval = FALSE }
prepped_penguin <- prep(softmaxkitchen, new_data = penguin_train)
prepped_penguin %>% juice() %>% get_condition(pen_length_mm) -> penguin_ped_len
prepped_penguin %>% juice() %>% get_condition(flipper_length_mm) -> penguin_flipper_len
penguin_ped_len
penguin_flipper_len
penguin_sex
prepped_penguin %>% bake(new_data = penguin_train) -> penguin_data
penguin_data %>% juice() -> penguin_data_j
penguin_data_j
penguin_pen_len <- penguin_data_j %>
```
